---
title: "Final Project" 
subtitle: "Stat 184"
author: "Tre Butler, Phil Saroka, Jessica Schick"
output: html_notebook
---

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())
# all packages used for the assignment
library(mosaicData)
library(mosaic)
library(DataComputing)
library(tidyverse)
library(forcats)
library(lubridate)
library(readxl)
```

Tre Butler, Phil Saroka, Jessica Schick
#Research Question:

How have outside factors like weather and time affected the crime rates of Boston?

#Data Sources:

##Where did you find them? (I don’t think we should just say Google)
- Boston crime rates data - https://www.kaggle.com/ankkur13/boston-crime-data
- Boston weather data - https://www.ncdc.noaa.gov/

##Who collected/maintains them?
- Boston crime rates data - AnkurJain formed the dataset on Kaggle, but the information to create the dataset comes from the Boston Police Department
- Boston weather data - The NOAA gathers it’s data from many sources including the World Meteorological Organization

##When & Why were they originally collected? 

A) Boston crime rates data - The dataset started in 2015 and was originally collected to find trends in crime statistics.
B) Boston weather data - The NOAA has kept weather statistics for many places for the past 30 years. They are collected for convenience to those who need to study trends in the weather.

##What does a case represent in each data source, and how many total cases are available?
Boston crime rates data - each case represent an Incident Number and there are a total of 260,760 cases available to use
Boston weather data - each case represent the Date of the Year and there are a total of 1,221 cases available to use

##What are some of the variables that you plan to use?
Boston crime rates data - Incident Number, Occurred on Date, Year, Month, Day of the Week, Hour, UCR Part, Street, Latitude, Longitude, Location, Offense Code
Boston weather data - Name, Latitude, Longitude, Date, Precipitation, Snow, Temperature Average
Plots, Summary Stats, and Observations

##TWO Data Sources (at least)
Primary data:  Boston Crime Data 
Other data sources: Boston Weather Data


## Joining datasets to create CrimeVsWeather Table
```{r}
CrimeVsWeather <-
  boston_crime_statistics %>%
  mutate("DATE" = ymd(as.Date(OCCURRED_ON_DATE))) %>%
  left_join(weather_data,"DATE")
```

## Using date verbs to create data for our graphic
```{r}
CrimeVsWeather <- 
  CrimeVsWeather %>%
  group_by(TAVG) %>%
  summarise(Total = n()) %>%
  arrange(desc(Total))
CrimeVsWeather
```

## Bar Graph Showing Avg Temperature Vs Crimes Committed
```{r}
CrimeVsWeather %>% 
  ggplot(aes(x=TAVG,y=Total )) +
  geom_bar(stat='identity',position='stack', width=.9) 
```
