---
title: "Final Project" 
subtitle: "Stat 184"
author: "Tre Butler, Phil Saroka, Jessica Schick"
output: html_notebook
---

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaicData)
library(mosaic)
library(DataComputing)
library(tidyverse)
library(forcats)
library(lubridate)
library(readxl)
```

### Research Question

How have outside factors like weather and time affected crime rates in Boston?

### Data Sources

1. Where did you find them?

Both data sets were found on Google.

2. Who collected/maintains them?

AnkurJain

https://www.kaggle.com/ankkur13/boston-crime-data

3. When & Why were they originally collected?

<<<<<<< HEAD
In 2018 mainly, to track crime stats in Boston
=======
In 2018 mainly, to track crime stats in Boston.
>>>>>>> 397617667f2212e2f5c357bfbf283504c5256eae

4. What does a case represent in each data source, and how many total cases are available?

An incident, over 300,000.

5. What are some of the variables that you plan to use?

### Plots, Summary Statistics, and Observations

### TWO Data Sources (at least)

#DATA

```{r}
#creating the datasets
boston_crime_statistics <- read_excel("crime.xlsx")
weather_data <- read_csv("weather data.csv")
weather_data <- weather_data%>%
  select(NAME,LATITUDE,LONGITUDE,DATE,PRCP, TAVG,SNOW)
```

```{r}
boston_crime_statistics
```
```{r}
weather_data
```
```{r}
all_stats <- cbind(boston_crime_statistics,weather_data)
```

```{r}
date_of_incident <- boston_crime_statistics%>%
  select(INCIDENT_NUMBER,OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)
loaction_of_incident <- boston_crime_statistics%>%
  select(INCIDENT_NUMBER,OCCURRED_ON_DATE,UCR_PART,STREET,Lat,Long, Location)%>%
  arrange(OCCURRED_ON_DATE)
```

```{r}
date_of_incident
```
```{r}
loaction_of_incident
```

```{r}
#date ranges
first_date<-boston_crime_statistics%>%
  select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  head(1)

last_date<-boston_crime_statistics%>%
  select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  tail(1)
first_date
last_date
```

```{r}
#number of day over which the data is recorded
last_date[1, "OCCURRED_ON_DATE"]-first_date[1, "OCCURRED_ON_DATE"]
```
```{r}
small_data <- boston_crime_statistics%>%
  group_by(OFFENSE_CODE)%>%
  summarise(total = n())
```
```{r}
ggplot(data=small_data,aes(x=OFFENSE_CODE,y=total))+geom_point()
```
