---
title: "Final Project" 
subtitle: "Stat 184"
author: "Tre Butler, Phil Saroka, Jessica Schick"
output: html_notebook
---

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaicData)
library(mosaic)
library(DataComputing)
library(tidyverse)
library(forcats)
library(lubridate)
library(readxl)
```

Tre Butler, Phil Saroka, Jessica Schick
#Research Question:

How have outside factors like weather and time affected the crime rates of Boston?

#Data Sources:

##Where did you find them? (I don’t think we should just say Google)
- Boston crime rates data - https://www.kaggle.com/ankkur13/boston-crime-data
- Boston weather data - https://www.ncdc.noaa.gov/

##Who collected/maintains them?
- Boston crime rates data - AnkurJain formed the dataset on Kaggle, but the information to create the dataset comes from the Boston Police Department
- Boston weather data - The NOAA gathers it’s data from many sources including the World Meteorological Organization

##When & Why were they originally collected? 

A) Boston crime rates data - The dataset started in 2015 and was originally collected to find trends in crime statistics.
B) Boston weather data - The NOAA has kept weather statistics for many places for the past 30 years. They are collected for convenience to those who need to study trends in the weather.

##What does a case represent in each data source, and how many total cases are available?
Boston crime rates data - each case represent an Incident Number and there are a total of 260,760 cases available to use
Boston weather data - each case represent the Date of the Year and there are a total of 1,221 cases available to use

##What are some of the variables that you plan to use?
Boston crime rates data - Incident Number, Occurred on Date, Year, Month, Day of the Week, Hour, UCR Part, Street, Latitude, Longitude, Location, Offense Code
Boston weather data - Name, Latitude, Longitude, Date, Precipitation, Snow, Temperature Average
Plots, Summary Stats, and Observations

##TWO Data Sources (at least)
Primary data:  Boston Crime Data 
Other data sources: Boston Weather Data

```{r}
#creating the datasets
boston_crime_statistics <- read_excel("crime.xlsx")
weather_data <- read_csv("weather data.csv")
weather_data <- weather_data%>%
  select(NAME,LATITUDE,LONGITUDE,DATE,PRCP, TAVG,SNOW)
```

```{r}
boston_crime_statistics
```
```{r}
weather_data
```
```{r}
date_of_incident <- boston_crime_statistics%>%
  select(INCIDENT_NUMBER,OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)
loaction_of_incident <- boston_crime_statistics%>%
  select(INCIDENT_NUMBER,OCCURRED_ON_DATE,UCR_PART,STREET,Lat,Long, Location)%>%
  arrange(OCCURRED_ON_DATE)
```

```{r}
date_of_incident
```
```{r}
loaction_of_incident
```

```{r}
#date ranges
first_date<-boston_crime_statistics%>%
  select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  head(1)

last_date<-boston_crime_statistics%>%
  select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  tail(1)
first_date
last_date
```

```{r}
#number of day over which the data is recorded
last_date[1, "OCCURRED_ON_DATE"]-first_date[1, "OCCURRED_ON_DATE"]
```
```{r}
small_data <- boston_crime_statistics%>%
  group_by(OFFENSE_CODE)%>%
  summarise(total = n())
```

##Create some informative plots and summary statistics 
```{r}
ggplot(data=small_data,aes(x=OFFENSE_CODE,y=total))+geom_point()
```
In this graph we can see that there are clusters of offense codes that are more common that others
