---
title: "Final Project" 
author: "Philip Saroka"
output: html_notebook
---
```{r}
#install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", 
#"ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata", 
#"rworldxtra","ggmap","digest","glue","cowplot"))
```
```{r}
#install.packages()
```
```{r warning=TRUE}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaicData)
library(mosaic)
library(DataComputing)
library(tidyverse)
library(forcats)
library(lubridate)
library(readxl)
library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(sp)
library(rworldmap)
library(rworldxtra)
library(leaflet)
library(data.table)
library(rgdal)
library(KernSmooth)
library(raster)
```

Tre Butler, Phil Saroka, Jessica Schick
#Research Question:

How have outside factors like weather and time affected the crime rates of Boston?

#Data Sources:

##Where did you find them? (I don’t think we should just say Google)
- Boston crime rates data - https://www.kaggle.com/ankkur13/boston-crime-data
- Boston weather data - Global Historical Climatology Network

##Who collected/maintains them?
- Boston crime rates data - AnkurJain formed the dataset on Kaggle, but the information to create the dataset comes from the Boston Police Department
- Boston weather data - The NOAA gathers it’s data from many sources including the World Meteorological Organization

##When & Why were they originally collected? 

A) Boston crime rates data - The dataset started in 2015 and was originally collected to find trends in crime statistics.
B) Boston weather data - The NOAA has kept weather statistics for many places for the past 30 years. They are collected for convenience to those who need to study trends in the weather.

##What does a case represent in each data source, and how many total cases are available?
Boston crime rates data - each case represent an Incident Number and there are a total of 260,760 cases available to use
Boston weather data - each case represent the Date of the Year and there are a total of 1,221 cases available to use

##What are some of the variables that you plan to use?
Boston crime rates data - Incident Number, Occurred on Date, Year, Month, Day of the Week, Hour, UCR Part, Street, Latitude, Longitude, Location, Offense Code
Boston weather data - Name, Latitude, Longitude, Date, Precipitation, Snow, Temperature Average
Plots, Summary Stats, and Observations

##Create some informative plots and summary statistics 

##Describe preliminary observations and intuition about the research question

##TWO Data Sources (at least)
Primary data:  Boston Crime Data 
Other data sources: Boston Weather Data









```{r message=FALSE, warning=FALSE}
#creating the datasets
boston_crime_statistics <- read_excel("crime.xlsx")
weather_data <- read_csv("weather data.csv")
weather_data <- weather_data%>%
  dplyr::select(NAME,LATITUDE,LONGITUDE,DATE,PRCP, TAVG,SNOW)
```

```{r}
boston_crime_statistics
```
```{r}
weather_data
```

```{r}
date_of_incident <- boston_crime_statistics%>%
  dplyr::select(INCIDENT_NUMBER,OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)
location_of_incident <- boston_crime_statistics%>%
  dplyr::select(INCIDENT_NUMBER,OCCURRED_ON_DATE,UCR_PART,STREET,Lat,Long, Location)%>%
  arrange(OCCURRED_ON_DATE)
```

```{r}
date_of_incident
```
```{r}
location_of_incident
```

```{r}
#date ranges
first_date<-boston_crime_statistics%>%
  dplyr::select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  head(1)

last_date<-boston_crime_statistics%>%
  dplyr::select(OCCURRED_ON_DATE,YEAR,MONTH,DAY_OF_WEEK,HOUR)%>%
  arrange(OCCURRED_ON_DATE)%>%
  tail(1)
first_date
last_date
```

```{r}
#number of day over which the data is recorded
last_date[1, "OCCURRED_ON_DATE"]-first_date[1, "OCCURRED_ON_DATE"]
```
```{r}
small_data <- boston_crime_statistics%>%
  group_by(OFFENSE_CODE)%>%
  summarise(total = n())
```
```{r}
ggplot(data=small_data,aes(x=OFFENSE_CODE,y=total))+geom_point()
```
In this graph we can see that there are clusters of offense codes that are more common that others

```{r}

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(40, 42), ylim = c(70, 72), expand = FALSE) +
    geom_density_2d(data=location_of_incident%>%head(10),aes(x=Long,y=Lat))
data(world.cities)


```

```{r}
location_of_incident
```
```{r}
lang<-location_of_incident%>%filter(Lat>40)%>%filter(Long<(-70))
lang
```
```{r}
lat_avg=mean(lang$Lat,na.rm = T)
lng_avg=mean(lang$Long,na.rm = T)
lat_avg
lng_avg
```


```{r}
ggplot(data=lang,aes(x=Long,y=Lat))+geom_density_2d()
```
```{r warning=FALSE}
setattr(lang, "class", c("data.table"))
kde <- bkde2D(lang[ , list(Long, Lat)],
              bandwidth=c(.0045, .0068), gridsize = c(100,100))

KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

palRaster <- colorNumeric("Spectral", domain = KernelDensityRaster@data@values, na.color = "transparent")

leaflet() %>% addTiles() %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .8) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Crime Heat Map")
  
```


